Hemodynamic responses of the Parahippocampal Place Area to spatial cues in a movie and its audio-description

### 2020-07-21 paper/main.tex
revision of introduction

\todo[inline]{Vorgabe: Intro + Discussion = main text ca. 3,000 words; einige
offene einige Dinge sind folgend in Kommentaren und nicht als ToDo gemarkt}

\todo[inline]{include figure providing schematic overview of the study and assay
design -> is overkill: Steht in den entsprechenden Veröffentlichungen \&
Paradigma ist mit "Film ab" wenig kompliziert}

@@

\section{Introduction} Cognitive neuroscientists use brain imaging methods like
blood oxygenation level-dependent functional magnetic resonance imaging (BOLD
fMRI) to map perceptual and cognitive brain processes (e.g. viewing of
landscapes, \citep{epstein1998ppa}; or theory of mind
\citep{spunt2014validating}) to functional brain areas or networks. Brain
mapping studies have traditionally employed small sets of carefully chosen,
simplified stimuli to strictly control experimental conditions and facilitate
functional isolation of brain areas. The disadvantage is that a limited number
of simplified stimuli does not resemble our real-life experience and lacks
external validity \citep{westfall2016fixing} as well as ecological validity
\citep{hasson2004intersubject}. Additionally, brain mapping studies usually
average results across at least 10-15 subjects to improve the
signal-to-noise-ratio (SNR). The disadvantage here is that the averaging
approach does not characterize brain functions at an individual level, a
prerequisite for the application of brain imaging methods in individual
diagnostics (cf. \citep{dubois2016building, eickhoff2020towards})

\todo[inline]{Retrosplenial Cortex kommt in Einleitung bisher  nicht vor; wird
aufgrund der Ergebnisse in AV und AO in Discussion diskutiert werden; cluster
sind in den Abbildungen nicht zu sehen; kommen aber in (den Tabellen) der
Results Section }

% What is the PPA?
A classic example of a functional, higher-level visual area is the
``parahippocampal place area'' (PPA) \citep{epstein1998ppa,
epstein1999parahippocampal}. The PPA is located in the posterior parahippocampal
gyrus at the boundary to the anterior lingual gyrus and medial fusiform gyrus,
both being part of the ventral visual stream \citep{mishkin1982contribution,
goodale1992separate}. Increased hemodynamic activity in the PPA correlates with
the perception of static pictures of landscapes or landmarks compared to
pictures of tools or faces.  According to the spatial layout hypothesis, the PPA
is involved in processing the surface geometry of a visual stimulus and possibly
in identifying scenes based on their spatial layout \citep{epstein2010reliable}.

% Literature Review PPA: Intro
At a group average level, results generalize to mental imagination of landscapes
\citep{ocraven2000mental} and haptic exploration of scenes constructed from LEGO
blocks \citep{wolbers2011modality}. One study that used spoken, place-related
sentences showed significantly decreased activation in the left and no
modulation in the right-hemispheric PPA \citep{aziz2008modulation}.

% Exemplary studies
% O'Craven: Mental imagery
In the study conducted by \citep{ocraven2000mental} participants viewed
alternating blocks of pictures showing famous faces and familiar places during
an initial experimental paradigm. In a subsequent paradigm, participants were
instructed to ``form a vivid mental image'' of the previously viewed pictures.
The PPA showed increased activation during imagination of places compared to
faces but the imagination tasks showed a smaller activation level compared to
the perceptual task.

@@

\todo[inline]{Er hat noch eine Connectivitäts-Analyse mit dem Occipital Cortex
gemacht. Im Vergleich zum visuellen Paradigma ist im haptischen Paradigma keine
Connectivität. Interpretation: Es kann keine mental imagery sein}

\todo[inline]{"The scene-related increase in coupling with the PPA was
significantly stronger in multiple clusters in occipital cortex under visual
than haptic stimulation (Figure 2B; Table S2)"}

@@

\todo[inline]{Ihre Method Section ist strange und es kommt bei ihr überall eh
    nur uneinheitliche Grütze heraus; Ich blicke nicht durch; was folgend kommt,
    wird nicht vollständig korrekt sein. So wie ich es verstanden habe, hat sie
    den Unterschied zwischen PPA und FFA angeguckt, aber diesen Unterschied im
Kontrast zu object-related Sentences (dortige Fig 3; lass uns kurz im Gespräch
drübergucken, ich habe die relvanten Abschnitte in der PDF markiert)}

To our knowledge only one study by \citep{aziz2008modulation} correlated
hemodynamic activity of the PPA with spoken sentences. Sentences described
generic or famous places (e.g. ``The Taj Mahal faces a long thin reflecting
pool''), faces  (e.g. ``Marilyn Monroe has a large square jaw''), or objects
(e.g. ``The television has a long antenna')'. Participants were instructed to
press a button whenever the sentence described an inaccurate or improbable fact.
Surprisingly, activation in the left, but not right, PPA was significantly
reduced when participants listened to place-related sentences compared to
listening to face-related sentences. Moreover, this effect was only observed in
sentences involving famous places.

\todo[inline]{Review von Aminoff (2013) kommt nur kurz als "competing hypothesis"; muss man dann natürlich in der Discussion diskutieren; was aber ggf. gar nicht notwendig ist, weil's darum gar bei uns gar nicht geht}

% Summary if literature review
In summary, studies suggest that the PPA does not exclusively respond to
visual(ly) spatial information. This is consistent with a competing hypotheses
stating that the PPA does not process the spatial layout of a scene specifically
but contextual associations in general (\citep{aminoff2006parahippocampal,
aminoff2013role}). It is still uncertain if auditory stimuli lead to inhibitory
or excitatory responses, or no change at all in the PPA.

\todo[inline]{aber Huth (2016) implizit; diskutiert PPA also solches jedoch
nicht und macht voxel-wise encoding, was - glaube ich - über
excitation/inhibition nichts aussagt -> Paper checken}

% Shortcomings of previous studies; lack of research
All reviewed studies used a just a small set of carefully designed, simplified
stimuli to strictly control experimental conditions. They further relied on
explicit judgment-based tasks to ensure that participants payed attention to
the stimuli. This raises the question how the PPA behaves under life-like conditions.

% Why naturalistic stimuli
Naturalistic stimuli like movies \citep{hasson2008neurocinematics,
sonkusare2019naturalistic} or narratives \citep{honey2012not,
lerner2011topographic, silbert2014coupled} offer complex, continuous stimuli
that better resemble our experience of a our dynamic environment.

% Literature review naturalistic stimuli: Einleitung
Two early studies using naturalistic stimuli suggest functional specialization
of brain areas is preserved during movie watching when many perceptual and
cognitive processes are triggered simultaneously \citep{bartels2004mapping, hasson2004intersubject}.

% Bartels (2004): Film + GLM (jedoch ohne PPA) und Hasson (2004)
\citep{bartels2004mapping} manually annotated time points that depict faces and
human bodies in a 22 minute clip of the movie \textit{Tomorrow Never Dies}
\citep{tomorrowneverdies}.
Employing a general linear model (GLM) to model hemodynamic activity, these time
points correlated with increased activity in the ``fusiform face area''
\citep{kanwisher1997ffa} and ``extrastriate body area''
\citep{downing2001bodyarea} respectively.

% Hasson (2004): Film + (reverse) ISC
Employing a reverse intersubject correlation, \citep{hasson2004intersubject}
used a 30 minute clip of the movie \textit{The Good, the Bad, and the Ugly}
\citep{goodbadugly} fMRI brain volumes showing the highest hemodynamic activity
in the PPA [across (five) participants] were matched to the temporally
corresponding movie frames. Results revealed that corresponding movie frames
depicted indoor and outdoor scenes including buildings and open fields.

% Summary naturalistic stimuli
Nevertheless, no study investigated how the PPA responds to spatial clues that are embedded in a complex, audio-only stimulus.
\todo[inline]{Huth (2016) checken}

\todo[inline]{The next paragraphs contain a bunch of methods already; reduce
this to talk about the general approach:

a) annotation of events

b) investigation of model quality,

c) cross-analysis of datasets

d) Possibly cite https://psyarxiv.com/sdbqv/ (DuPre (2019). Nature abhors a
paywall) and declare that this study is a concrete realization of that promise}

% what we did data source
In this study, we analyze functional data of the same 14 participants provided
by the studyforrest dataset \href{http://www.studyforrest.org}{studyforrest.org}; \citep{hanke2014audiomovie, hanke2016simultaneous}).
% PPA + movie + audio-description
We annotated cuts in the movie ``Forrest Gump'', and words spoken by an
additional narrator in the movie's audio-description (i.e. the movie's
audio-only variant) to exploratorily investigate how the parahippocampal gyrus
behaves during lifelike stimulation.
% model-driven analysis
In conformity to the analyses performed in the reviewed studies,  we also
employed a standard multi-level, voxel-wise general linear model
(GLM).\todo{why? in order to\dots}
% focus on audio-description
Encouraged by \citep{aziz2008modulation} and \citep{aminoff2013role}, we focus
on the audio-description as a naturalistic, auditory-only paradigm free from any task to localize a ``visual area'' in individual study participants.
% comparison to dedicated localizer
Finally, we compare our results to results of a dedicated visual localizer
experiment done with the same 14 participants \citep{sengupta2016extension}.

@@

\todo[inline]{Specific data outputs should be explicitly referenced via data
citation (see Data Records and Data Citations, below)}
\todo[inline]{order of subsections possibly wrong}
\todo[inline]{parts from original studies still detailed}

% intro
We used data from three publicly available datasets
(\href{http://www.studyforrest.org}{studyforrest.org}) which have already been
used by other research groups for independent research questions.
% used studies
\todo{this is just loveless listing so far but it's 'complete'} The same
subjects were
% AO
a) listening to the audio-description (AO study; \citep{hanke2014audiomovie}) of
the movie ``Forrest Gump'', a dataset already used by \citep{hu2017decoding,
nguyen2016integration},
% AV
b) watching the actual audio-visual movie (AV study;
\citep{hanke2016simultaneous}, a dataset already used by
\citep{ben2018hippocampal},
% VIS
c) participating in a dedicated six-category block-design visual localizer (VIS
study) \citep{sengupta2016extension}, a dataset already used by
\citep{jiahui2019predicting}.
% see corresponding papers for details
An exhaustive description of the participants, stimulus creation, procedure,
stimulation setup, and fMRI acquisition can be found in the corresponding
publications \citep{hanke2014audiomovie, hanke2016simultaneous,
sengupta2016extension}. Following is a summary of most important aspects.

>>>>

\todo[inline]{Intro + Discussion = main text; ca. 3,000 words; some minor
    ``issues'' are in comments and not labeled as to do's}

@@

\section{Introduction}

% brain mapping via fMRI
Cognitive neuroscientists use brain imaging methods like blood oxygenation
level-dependent functional magnetic resonance imaging (BOLD fMRI) to map
perceptual and cognitive brain processes (e.g. viewing of landscapes,
\citep{epstein1998ppa}; or theory of mind \citep{spunt2014validating}) to
functional brain areas or networks.
% simplified stimuli
Brain mapping studies have traditionally employed small sets of carefully
chosen, simplified stimuli to strictly control experimental conditions.
% disadvantage 1: validity
Their disadvantage is that a limited number of simplified stimuli does not
resemble our real-life experience and thus lack external validity
\citep{westfall2016fixing} as well as ecological validity
\citep{hasson2004intersubject}.
% disadvantage 2: averaging for group results
Additionally, brain mapping studies usually average results across at least
10-15 subjects to improve the signal-to-noise-ratio (SNR). The disadvantage here
is that the averaging approach does not characterize brain functions at an
individual level, a prerequisite for the application of brain imaging methods in
individual diagnostics (cf. \citep{dubois2016building, eickhoff2020towards}).

\todo[inline]{RSC is not mentioned in the introduction but should (probably)
shortly discussed in the, well discussion; RSC shows up in group results but
mixed results in individual subjects}

% What is the PPA?
A classic example of a functional, higher-level visual area is the
``parahippocampal place area'' (PPA) \citep{epstein1998ppa,
epstein1999parahippocampal}.
% anatomical location
The PPA is located in the posterior parahippocampal gyrus at the boundary to the
anterior lingual gyrus and medial fusiform gyrus, both being part of the ventral
visual stream \citep{mishkin1982contribution, goodale1992separate}.
% neural correlate of scene perception
Increased hemodynamic activity in the PPA correlates with the perception of
static pictures of landscapes or landmarks compared to pictures of tools or
faces.
% spatial layout hypothesis
According to the spatial layout hypothesis, the PPA is involved in
processing the surface geometry of a visual stimulus and possibly in identifying
scenes based on their spatial layout \citep{epstein2010reliable}.

% literature review overview 1: imagination & haptic exploration
At a group average level, results generalize to mental imagination of landscapes
\citep{ocraven2000mental} and haptic exploration of scenes constructed from LEGO
blocks \citep{wolbers2011modality}.
% literature review overview 2: spoken sentences
One study that used spoken, place-related sentences showed significantly
decreased activation in the left and no modulation in the right-hemispheric PPA
\citep{aziz2008modulation}.

% literature review in detail
% O'Craven: Mental imagery
In the study conducted by \citep{ocraven2000mental} participants viewed
alternating blocks of pictures showing famous faces and familiar places during
an initial experimental paradigm. In a subsequent paradigm, participants were
instructed to ``form a vivid mental image'' of the previously viewed pictures.
The PPA showed increased activation during imagination of places compared to
faces but the imagination tasks showed a smaller activation level compared to
the perceptual task.

@@

\todo[inline]{Er hat noch eine Connectivitäts-Analyse mit dem Occipital Cortex
    gemacht. Im Vergleich zum visuellen Paradigma ist im haptischen Paradigma
    keine Connectivität. Interpretation: Es kann keine mental imagery sein;
    ``The scene-related increase in coupling with the PPA was significantly
    stronger in multiple clusters in occipital cortex under visual than haptic
    stimulation (Figure 2B; Table S2)''}

@@

\todo[inline]{lass uns kurz im Gespräch über Aziz Paper gucken; die relvanten
    Abschnitte sind in der PDF markiert. Ihre Method Section ist ``strange'' und
    im Grunde kommt überall nur uneinheitliche Grütze heraus; ich blicke nicht
    durch, weshalb folgendes nicht vollständig korrekt sein wird;so wie ich es
    verstanden habe, hat sie den Unterschied zwischen PPA und FFA angeguckt,
aber diesen Unterschied im Kontrast zu object-related Sentences (dortige Fig 3;
)}

% Aziz' stimuli
To our knowledge only one study by \citep{aziz2008modulation} correlated
hemodynamic activity of the PPA with spoken sentences. Sentences described
generic or famous places (e.g. ``The Taj Mahal faces a long thin reflecting
pool''), faces  (``Marilyn Monroe has a large square jaw''), or objects (``The
television has a long antenna'').
% Aziz' tasks
Participants were instructed to press a button whenever the sentence described
an inaccurate or improbable fact.
% Aziz' results
Activation in the left, but not right, PPA was significantly reduced when
participants listened to place-related sentences compared to listening to
face-related sentences. Moreover, this effect was only observed in sentences
involving famous places.

\todo[inline]{I deleted sentences regarding Aminoff's
    (\citep{aminoff2006parahippocampal, aminoff2013role}) competing ``contextual
associations hypothesis'' because our paradigm does not touch that issue.}

% summary if literature review
In summary, studies suggest that the PPA does not exclusively respond to
visually presented spatial information. Nevertheless, the question how the
PPA responds to auditory stimuli is still open.

\todo[inline]{but Huth et al. (2016) implicitly; but they do not discuss the
PPA; moreover, they do voxel-wise encoding which doesn't say anything about
excitation/inhibition}

% shortcomings of previous studies: simplified stimuli
All reviewed studies used a small set of carefully designed, simplified stimuli
to strictly control experimental conditions.
% shortcomings of previous studies: task-based paradigm
They further relied on explicit judgment-based tasks to ensure that participants
payed attention to the stimuli.
% second question: life-like conditions
This raises the second question which is how the PPA behaves under life-like
conditions.
% naturalistic stimuli
Naturalistic stimuli like movies \citep{hasson2008neurocinematics,
sonkusare2019naturalistic} or narratives \citep{honey2012not,
lerner2011topographic, silbert2014coupled} offer complex, continuous stimuli
that better resemble our experience of a our dynamic environment.

% naturalistic stimuli review: intro
Two early studies using naturalistic stimuli suggest functional specialization
of brain areas is preserved during movie watching when many perceptual and
cognitive processes are triggered simultaneously \citep{bartels2004mapping, hasson2004intersubject}.
% Bartels (2004): Film + GLM (but surprisingly did not annotate landscapes)
\citep{bartels2004mapping} manually annotated time points that depict faces and
human bodies in a 22 minute clip of the movie \textit{Tomorrow Never Dies}
\citep{tomorrowneverdies}.
% Bartels: results
Employing a general linear model (GLM) to model hemodynamic activity, these time
points correlated with increased activity in the ``fusiform face area''
\citep{kanwisher1997ffa} and ``extrastriate body area''
\citep{downing2001bodyarea} respectively.
% Hasson (2004): Film + (reverse) ISC
Employing a reverse intersubject correlation, \citep{hasson2004intersubject}
used a 30 minute clip of the movie \textit{The Good, the Bad, and the Ugly}
\citep{goodbadugly}.
% Hasson: method
fMRI brain volumes showing the highest hemodynamic activity
in the PPA [across (five) participants] were matched to the temporally
corresponding movie frames.
% Hasson: results
Results revealed that corresponding movie frames
depicted indoor and outdoor scenes including buildings and open fields.

% summary naturalistic stimuli
Nevertheless, both studies used an audio-visual stimulus and thus do not address
the question how the PPA responds to spatial clues that are embedded in a
complex, audio-only stimulus.

\todo[inline]{mih: Possibly cite https://psyarxiv.com/sdbqv/ (DuPre (2019).
Nature abhors a paywall) and declare that this study is a concrete realization
of that promise}

% what we did: data source
In this study, we analyze functional data of the same 14 participants being part
of the studyforrest dataset
(\href{http://www.studyforrest.org}{studyforrest.org};
\citep{hanke2014audiomovie, hanke2016simultaneous}).
% PPA + movie + audio-description
We annotated cuts in the movie ``Forrest Gump'' \citep{ForrestGumpMovie}, and
words spoken by an additional narrator in the movie's audio-description (i.e.
the movie's audio-only variant; \citep{ForrestGumpGermanAD}) to exploratorily
investigate how the parahippocampal gyrus behaves during movie watching and
listening to an ``audiobook''.
% focus on audio-description
Encouraged by \citep{aziz2008modulation}, we also employ a traditional GLM, and
focus on the audio-description as a naturalistic, auditory-only paradigm free
from any task to localize a ``visual area'' in individual study participants.
% comparison to dedicated localizer
We also compare our results to results of a dedicated visual localizer
experiment done with the same participants \citep{sengupta2016extension}.

@@

\todo[inline]{Specific data outputs should be explicitly referenced via data
citation (see Data Records and Data Citations, below)}
\todo[inline]{current order of subsections might be ``non-canonical''}
\todo[inline]{parts from original studies still detailed; feel free to delete
sentences}

% intro
We used data from three publicly available datasets
(\href{http://www.studyforrest.org}{studyforrest.org}) which have already been
used by other research groups for independent research questions.
% used studies
\todo{this is just loveless listing so far but it's 'complete'} The same
subjects were
% AO
a) listening to the audio-description (AO study; \citep{hanke2014audiomovie}) of
the movie ``Forrest Gump'', a dataset already used by \citep{hu2017decoding,
nguyen2016integration},
% AV
b) watching the actual audio-visual movie (AV study;
\citep{hanke2016simultaneous}), a dataset already used by
\citep{ben2018hippocampal},
% VIS
c) participating in a dedicated six-category block-design visual localizer (VIS
study; \citep{sengupta2016extension}), a dataset already used by
\citep{jiahui2019predicting}.
% see corresponding papers for details
An exhaustive description of the participants, stimulus creation, procedure,
stimulation setup, and fMRI acquisition can be found in the corresponding
publications. Following is a summary of most important aspects.
